Flaring star-planet interactions in Kepler and TESS

### 2023-06-21 src/tex/ms.tex
[showyourwork] overleaf sync

\subsection{Magnetic star-planet interactions}
\label{sec:intro:mspi}
Magnetic star-planet interactions can occur if the planet is orbiting within the Alfv\'en radius of the star at least part of the time~\citep{preusse2006magnetic, cohen2011dynamics}. The Alfv\'en radius is the radius at which the stellar wind velocity, which increases with increasing distance from the star, exceeds the Alfv\'en velocity of the magnetized plasma. Beyond this radius, this plasma is disconnected from the star. Alfv\'en waves set off by the planet within the magnetic field they cross cannot propagate back to the star if the wind carries the plasma away faster than these waves can travel. In the sub-Alfv\'enic regime, the wave can reach the star, and deposit its energy in its atmosphere. \citet{lanza2012starplanet, lanza2018closeby}, and \citet{zarka2007plasma} and \citet{saur2013magnetic} suggest different mechanisms for when and how the energy transport and dissipation takes place, and how much energy can be transferred and dissipated at all. They all lack observational constraints from large samples for calibration, which we aim to provide in this work.

@@

The absolute power of flaring star-planet interactions is difficult to constrain. We do not know how energetic planet-induced flares can be. The estimates for dissipated energy~\citep{lanza2018closeby} are in the ballpark of the regime of superflares observed by Kepler and TESS. However, on the one hand, the planet might be triggering flares prematurely, which would be of lower energy~\citep{loyd2023flares}, and end up below our detection threshold. On the other hand, the energy need not be dissipated continuously, particularly in the stretch-and-break mechanism proposed by~\citep{lanza2012starplanet}, so an individual flare's energy might be higher. We also do not know whether flares triggered by interactions with a companion would have different properties than intrinsic flares. We will assume that, since they would be caused by magnetic reconnection triggered by the planet in the stellar corona, they should still have the same physics as intrinsic flares.

@@

\section{Data}
\label{sec:data}
In this study, we aimed at compiling the largest possible sample to search for flaring star-planet interactions. We took the entire Planetary Systems Composite Parameters Table (PSCP\footnote{ \url{https://exoplanetarchive.ipac.caltech.edu/cgi-bin/TblView/nph-tblView?app=ExoTbls&config=PSCompPars}}, Section~\ref{sec:data:sps}) from the NASA Exoplanet Archive, as of July 2022, and scanned the systems for available Kepler and TESS light curves~(Section~\ref{sec:data:photometry}). We searched a total of 1811 systems and over 7200 light curves from both missions for flares. To calculate the amount of excess flaring from star-planet interaction, we further required the orbital periods of each system~(Section~\ref{sec:data:orbitalperiod}). To compare our measurements to theory, we estimated the expected power of interaction for each system with at least three flares. For this, we required further system parameters -- distance between planet and star, stellar rotation period, planetary radius, and stellar luminosity~(Sections~\ref{sec:data:a}-\ref{sec:data:lum}), which we mostly took from the literature. We began by compiling the table of system parameters with the PSCP, but verified, supplemented and updated the table with more up-to-date results. The parameters for the systems in the final sample of 25 systems are shown in Table~\ref{tab:maintable_lit}.

@@

\section{Methods}
\label{sec:methods}
We measure flaring star-planet interactions (SPI) as the presence of excess flares triggered by the orbiting planet. In the absence of flaring SPI, flare peak times will be distributed randomly in orbital phase. In the presence of flaring SPI, we measure a phase dependent deviation from this randomness.

The main data for this analysis are flare times. To obtain them, we gather the Kepler and TESS light curves for all star-planet systems, remove rotational variability trends, search the de-trended light curves for flares, and convert the flare times to orbital phases. We then perform a customized Anderson-Darling test on the flare peak time distribution, which yields a $p$-value for the significance of the SPI signal. Finally, we compare these results to the theoretically expected SPI power $P_{spi}$ in each system.

@@

\subsection{Custom Anderson-Darling test}
\label{sec:methods:adtest}
For each star-planet system with three or more flares in its light curves, we test for deviations from a random distribution of flares with orbital phase. We use the same customized Anderson-Darling test as in \cite{ilin2022searching}, which is more sensitive to deviations at both ends of the distribution than the widely used Kolmogorov-Smirnov test~\citep{feigelson2012statistical}. In brief, we first take the number of flares observed in the TESS and Kepler data to calculate a base flare rate for each light curve. We then calculate how often each orbital phase has been observed, also for each light curve separately, to ensure that the different detection thresholds for TESS and Kepler do not bias the base rate. With the phase coverage and base flare rate combined, we can tell how many flares we would expect to see in any given phase bin if flares were randomly distributed. We then aggregate this number of expected flares per time bin into one expected distribution by summing over all available light curves. As a last step, we compare the expected distribution to the observed one using an Anderson-Darling test. We repeat this test with different phase offsets to account for potential biases in sensitivity of the test at different phases.

@@

\subsection{Power of star-planet interaction}
\label{sec:methods:pspi}
We consider two theories for the mechanism behind flaring star-planet interactions, the stretch-and-break and the Alfv\'en wing mechanisms. For both of them, we include the case of a magnetized and an unmagnetized planet, for a total of four estimates of the expected power of magnetic star-planet interaction.

First, we estimate the power of star-planet interaction generated by the stretch-and-break mechanism using the scaling relations in~\cite{lanza2012starplanet} (their Eqn. 45), where $P_{spi,sb}$ is proportional to

\begin{equation}
    P_{spi,sb} \sim B_p^{2/3} R_p^2 B_*^{4/3} R_*^4 v_{rel} a^{-4} .
\end{equation}

@@

\begin{equation}
    P_{spi,sb0} \sim R_p^2 R_*^4 B_*^2 v_{rel}    a^{-4} 
\end{equation}

Third, we estimate the power of star-planet interaction generated by the Alfv\'en wing mechanism ~\citep{saur2013magnetic,kavanagh2022radio}, combining Eqns.~8~and~11 in \citet{kavanagh2022radio} to arrive at

\begin{equation}
    P_{spi,aw} \sim B_p^{2/3} R_{p}^2    B_*^{1/3}  v_{rel}^2 a^{-2}  
\end{equation}
where we replace the stellar wind density $\rho_W$ with its proportionality to $a^{-2}$, and stellar wind magnetic field at the planetary radius $B_W$ with the surface field strength $B_* a^{-3}$, which result in the total $a^{-2}$ factor in the above equation.

@@

\begin{equation}
    P_{spi,aw0} \sim R_p^2  B_*  v_{rel}^2  a^{-4}
\end{equation}

Lacking $B_p$, and various additional factors that appear in $P_{spi,sb},P_{spi,sb0},P_{spi,aw}$, and  $P_{spi,aw0}$, such as the angle between the relative velocity and magnetic field of the stellar wind~\citep{kavanagh2022radio}, or the force-free parameter~\citep{lanza2012starplanet}, we restrict ourselves to differential estimates, and normalize all $P_{spi}$ to AU Mic.

@@

Our goal was to measure magnetic star-planet interaction as the statistical clustering of flares in phase with the innermost planet's orbit. We searched all star-planet systems that were observed with Kepler and TESS at 1 or 2 min cadence, respectively, for flares. For the star-planet systems in the resulting flare catalog~(Section~\ref{sec:results:catalog}), we confirmed that the orbital phase could be known well enough for the entire observing baseline of the system~(Section~\ref{sec:results:coherence}). We calculated the orbital phases of each flare, and the flare phase distributions for each system~(Section \ref{sec:results:phasedist}). We then applied the custom Anderson-Darling test to assess how much each distribution was different from randomly distributed intrinsic flaring. Comparing the test results with the expected power of magnetic star-planet interaction, we found that the amount of excess flaring tentatively increases with the expected power in all considered scenarios~(Section~\ref{sec:results:spi}). However, not all systems with high expected power show signs of flaring SPI, which creates two branches in the distribution of test results. Finally, in Section~\ref{sec:results:individualstars}, we provide additional context for the most interesting systems in our sample, and explain why we had to exclude others from the analysis.

@@

Figure~\ref{fig:adtest_bp} shows that in every scenario we considered (Alfv\'en wing or stretch-and-break mechanism, with magnetized or unmagnetized planets), there are two branches on the high end of expected powers. On one branch, the significance of the measured interaction increases with the expected power $P_{spi}$. On the other branch, no interaction is measured regardless of $P_{spi}$.

Table~\ref{tab:maintable_der} lists the $p$-values for each star-planet system. There is no star-planet system for which we find a $>3\sigma$ ($p<0.0027$) signal of star-planet interaction in our sample. However, overall, the significance of the deviation increases with higher expected power $P_{spi}$ of interaction, where additional subscripts denote different scenarios introduced in Section~\ref{sec:methods:pspi}. Table~\ref{tab:maintable_der} also lists the derived parameters required in these scaling laws, i.e., Rossby number $R$o, surface-average magnetic field strength $B$, and relative velocity between the planet and a co-rotating magnetic field $v_{\rm rel}$.

@@

\begin{table*}
\footnotesize
\movetableright=-20mm
\script{paper_main_table.py}
\caption{Flaring star-planet interaction. $Ro$, $B$, and $v_{rel}$, are derived from literature values (Table~\ref{tab:maintable_lit}). $P_{spi,xx}$ stands for the power of stretch-and-break ($sb$) and Alfv\'en wing ($aw$) interaction mechanisms, assuming the planet has a magnetic field strength of $1\,$G. $P_{spi,xx0}$ is the same, but assuming an unmagnetized planet. All powers are normalized to AU Mic. The $p$-value of the Anderson-Darling test is lower when the system shows more flares periodic with the planetary orbit.}
\input{output/table_der_vals.tex}
    \label{tab:maintable_der}

@@

\begin{figure*}[ht!]
    \script{paper_adtest_vs_value_scatterplots.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/PAPER_ADtest_bg.png}
        \caption{
            Expected power of SPI vs. AD test results, assuming four different scenarios, color-coded by stellar surface field strength. \textbf{Top panels}: stretch-and-break scenario. \textbf{Bottom panels}: Alfv\'en wing scenario. \textbf{Left panels}: assuming a magnetized planet. \textbf{Right panels}: assuming an unmagnetized planet. While the distribution of $p$-values is consistent with no interaction, all scenarios indicate lower $p$-values only for high expected powers of interaction. See Fig.~\ref{fig:adtest_ro} in the Appendix for the same figure, color-coded by Rossby number.
        }
        \label{fig:adtest_bp}
    \end{centering}
\end{figure*}

\subsection{Individual stars}
\label{sec:results:individualstars}
The properties of star-planet systems in this work are diverse, including both very fast and very slowly rotating host stars; systems with super-Earths, Neptunes, and Hot Jupiters; and spectral types covering the lower main sequence from mid-F to late M.

For our analysis, it is important to know that the observed flares indeed occurred on the planet host star. We therefore excluded some systems due to documented contamination from nearby (bound or co-moving) companions~(Section~\ref{sec:results:individualstars:excluded}). One exception to this rule is Kepler-411, a binary system for which the flare contributions from each component could be separated, but which was excluded for other reasons~(Section~\ref{sec:results:individualstars:kep411}). We also dropped GJ 1061 from the sample because its rotation period was unconstrained. Another uncertainty in our analysis is the possibility of additional planets at even shorter orbits than the currently known innermost planet, such as might be the case for Proxima Cen~(Section~\ref{sec:results:individualstars:proxima}).

The bulk of the systems in our sample are neither expected to show high power of magnetic star-planet interaction, nor do they show any deviation from intrinsic stellar flaring. In the subsample of stars with high expected power, of the order of AU Mic's and above, we first take a closer look at those that seem to follow an increase in measured flaring star-planet interaction with increasing expected power $P_{spi}$, that is AU Mic itself, K2-25, TOI-540, and HIP 67522~(Sections \ref{sec:results:individualstars:aumic} and \ref{sec:results:individualstars:hip67522}). Then we consider those systems where high power was expected in all scenarios, but not measured, i.e., TAP~26 and KOI-12~(Sections \ref{sec:results:individualstars:tap26} and \ref{sec:results:individualstars:koi12}).
\subsubsection{Excluded targets}
\label{sec:results:individualstars:excluded}

@@

\subsubsection{Prominent systems}
\label{sec:prominensystems}
\paragraph{Proxima Cen}
\label{sec:results:individualstars:proxima}
Although Proxima Cen is part of a triple stellar system, we treat it a single star in this work because of the large angular separation of over 2 deg to its companions, $\alpha$ Cen AB. We do not detect a deviation from intrinsic flaring on Proxima Cen, consistent with numerical models that place Proxima Cen b well outside the sub-Alfv\'enic zone~\citep{gomez2020earthlike, kavanagh2021planetinduced, garraffo2022revisiting}. However, this might not entirely exclude Proxima Cen from the search for flaring star-planet interaction. The tentatively detected Proxima Cen d, a planet further in, at 0.029 au, or 5 day orbital period~\citep{faria2022candidate, artigau2022linebyline} could still cause flaring star-planet interactions. Depending on the phase of the activity cycle, Proxima Cen d may be in the sub-Alf\'enic regime during Proxima Cen's activity maximum~\citep{gomez2020earthlike}. However, the orbital period of the tentative planet is so uncertain at this point that its coherence time~(Section~\ref{sec:results:coherence}) is shorter than the roughly 2yr observing baseline of Proxima Cen, preventing us from accurately measuring the flare phases.

@@

AU Mic is a 16-29 Myr old pre-main sequence M0-M1 dwarf with a strong magnetic field of about $3010\pm220\,$G obtained from Zeeman broadening measurements~\citep{reiners2022magnetism}. The innermost Neptune-sized planet, AU Mic b, could be both inside the sub-Alfv\'enic zone if the star's mass loss rate is relatively low at about 30 times the solar value or lower~\citep{alvarado-gomez2022simulating}. If the mass loss rate is high, exceeding the solar value by a factor of several hundreds, AU Mic b could orbit in the super-Alfv\'enic zone, and become unable to experience any planet-induced flaring. Recently, \citet{klein2022one} found tentative periodicity with the orbital period of AU Mic b in the chromospheric He I D line, which increases if the contribution from flares is included in the calculation. Overall, our marginal signal of flaring star-planet interaction in the AU Mic system supports the idea that this young, magnetically active M dwarf system with a close-in Neptune could exhibit planet-induced flaring.
For a detailed discussion of AU Mic and its flaring star-planet interaction, we refer to~\cite{ilin2022searching}, where we also estimate that an additional 50–100 days of TESS-like monitoring of AU Mic would yield a $3\sigma$ detection if the marginal signal in our data is real.

K2-25 is a multiplanet system around a fast rotating ($P_{\rm rot}<2\,$d) mid-M dwarf in the Hyades open cluster~\citep[600-800 Myr, ][]{stefansson2020habitable}. Both the expected power of interaction, and the measured deviation, are remarkably similar to the AU Mic system. Interestingly, K2-25 b is in a presumably similar environment to GJ 436 b, but does not show the atmospheric escape in Ly$\alpha$ that the latter famously exhibits, which might be due to high speed stellar winds that suppress atmospheric escape~\citep{rockcliffe2021lya, carolan2020dichotomy, vidotto2020stellar}. Further investigation of magnetic star-planet interactions, flaring or otherwise, in this system could add an important constraint on K2-25's wind properties by clarifying whether the planet orbits in- or outside the Alfv\'en radius, especially accounting for its moderately high eccentricity~(see also Section~\ref{sec:discussion:eccentricity}).

@@

\paragraph{TAP 26}
\label{sec:results:individualstars:tap26}
TAP 26 is a weak-line T-Tauri star with a strong magnetic field~\citep{yu2017hot}. \citet{lanza2018closeby} estimate that the system can release more energy in magnetic star-planet interaction than other  systems with Hot Jupiters, such as HD 179949, that has previously been detected with chromospheric variability in phase with the planet's orbit~\citep{shkolnik2008nature}. However, for TAP 26 b, the orbital period might not be well-constrained. \citet{yu2017hot} apply several methods to derive $P_{\rm orb}$ from the radial velocity data: While the $10.8\,$d orbit we adopted is the most likely according to \citet{yu2017hot}, a $9.0\,$d orbit is also likely, and some of their applied methods favored a $13.4\,$d period. The absence of interaction could hence be a consequence of uncertain orbital period. However, intermittent interactions and the viewing geometry of TAP 26 could also explain the absence of phase-correlated flaring~(see Sections~\ref{sec:discussion:intermittency} and \ref{sec:discussion:viewing}).

\paragraph{KOI-12}
\label{sec:results:individualstars:koi12}
KOI-12, also known as Kepler-448, is a two-planet system with an eccentric outer planet KOI-12 c, and an inner Warm Jupiter planet KOI-12 b in a $17.8\,$d orbit~\citep{masuda2017eccentric}. With respect to rotation rate, stellar and planetary radius, the KOI-12 system is similar to HIP~67522 and TAP~26 (although for the those young stars the radii are large because the stars are still contracting onto the main sequence). In constrast, KOI-12 is $1.4\pm0.3\,$Gyr old~\citep{bourrier2015sophie}, an F5 subgiant~\citep{frasca2016activity}, which is already evolving off the main sequence. Yet, despite its high expected power of interaction, we detect no excess flaring in phase with KOI-12 b. The reason could be that KOI-12 b is in fact super-Alfv\'enic due to the relatively wide orbit compared to TAP 26 and HIP 67522. However, our estimate of the Alf\'en radius place KOI-12 b in the sub-Alfv\'enic regime~(see Fig.~\ref{fig:as}), although the scaling law applied there may be inaccurate for post-main sequence stars. Intermittent interaction might play a role here as well, and could be favored if future observations capture episodes during which the interaction is 'on'.

@@

Magnetic star-planet interaction can only take place if the planet spends at least a part of its orbit in the sub-Alfv\'enic regime~(see Section~\ref{sec:intro:mspi}). We estimated the average Alfv\'en surface size ($\overline{AS}_{R}$) using the power law established by \citet{Chebly2023}:

@@

The third model uses the gravitational perturbation as a proxy for the interaction~\citep{cuntz2000stellar}, originally derived for star-planet interactions.  Their Eqn. 1 proposes a gravitational perturbation proportional to the mass ratio between planet and star, and $a^{-3}$.

>>>>

\subsection{Magnetic star-planet interactions}
\label{sec:intro:mspi}
Magnetic SPI can occur if the planet is orbiting within the Alfv\'en radius of the star at least part of the time~\citep{preusse2006magnetic, cohen2011dynamics}. The Alfv\'en radius is the radius at which the stellar wind velocity, which increases with increasing distance from the star, exceeds the Alfv\'en velocity of the magnetized plasma. Beyond this radius, this plasma is disconnected from the star. Alfv\'en waves set off by the planet within the magnetic field they cross cannot propagate back to the star if the wind carries the plasma away faster than these waves can travel. In the sub-Alfv\'enic regime, the wave can reach the star, and deposit its energy in its atmosphere. \citet{lanza2012starplanet, lanza2018closeby}, and \citet{zarka2007plasma} and \citet{saur2013magnetic} suggest different mechanisms for when and how the energy transport and dissipation takes place, and how much energy can be transferred and dissipated at all. They all lack observational constraints from large samples for calibration, which we aim to provide in this work.

@@

The absolute power of flaring SPI is difficult to constrain. We do not know how energetic planet-induced flares can be. The estimates for dissipated energy~\citep{lanza2018closeby} are in the ballpark of the regime of superflares observed by Kepler and TESS. However, on the one hand, the planet might be triggering flares prematurely, which would be of lower energy~\citep{loyd2023flares}, and end up below our detection threshold. On the other hand, the energy need not be dissipated continuously, particularly in the stretch-and-break mechanism proposed by~\citep{lanza2012starplanet}, so an individual flare's energy might be higher. We also do not know whether flares triggered by interactions with a companion would have different properties than intrinsic flares. We will assume that, since they would be caused by magnetic reconnection triggered by the planet in the stellar corona, they should still have the same physics as intrinsic flares.

@@

\section{Data}
\label{sec:data}
In this study, we aimed at compiling the largest possible sample to search for flaring SPI. We took the entire Planetary Systems Composite Parameters Table (PSCP\footnote{ \url{https://exoplanetarchive.ipac.caltech.edu/cgi-bin/TblView/nph-tblView?app=ExoTbls&config=PSCompPars}}, Section~\ref{sec:data:sps}) from the NASA Exoplanet Archive, as of July 2022, and scanned the systems for available Kepler and TESS light curves~(Section~\ref{sec:data:photometry}). We searched a total of 1811 systems and over 7200 light curves from both missions for flares. To calculate the amount of excess flaring from SPI, we further required the orbital periods of each system~(Section~\ref{sec:data:orbitalperiod}). To compare our measurements to theory, we estimated the expected power of interaction for each system with at least three flares. For this, we required further system parameters -- distance between planet and star, stellar rotation period, planetary radius, and stellar luminosity~(Sections~\ref{sec:data:a}-\ref{sec:data:lum}), which we mostly took from the literature. We began by compiling the table of system parameters with the PSCP, but verified, supplemented and updated the table with more up-to-date results. The parameters for the systems in the final sample of 25 systems are shown in Table~\ref{tab:maintable_lit}.

@@

\section{Methods}
\label{sec:methods}
We measure flaring SPI as the presence of excess flares triggered by the orbiting planet. In the absence of flaring SPI, flare peak times will be distributed randomly in orbital phase. In the presence of flaring SPI, we measure a phase dependent deviation from this randomness.

The main data for this analysis are flare times. To obtain them, we gather the Kepler and TESS light curves for all star-planet systems, remove rotational variability trends, search the de-trended light curves for flares, and convert the flare times to orbital phases. We then perform a customized Anderson-Darling test on the flare peak time distribution, which yields a $p$-value for the significance of the SPI signal. Finally, we compare these results to the theoretically expected SPI power $P_{\rm spi}$ in each system.

@@

\subsection{Custom Anderson-Darling test}
\label{sec:methods:adtest}
For each star-planet system with three or more flares in its light curves, we test for deviations from a random distribution of flares with orbital phase. We use the same customized Anderson-Darling test as in \cite{ilin2022searching}, which is more sensitive to deviations at both ends of the distribution than the widely used Kolmogorov-Smirnov test~\citep{feigelson2012modern}. In brief, we first take the number of flares observed in the TESS and Kepler data to calculate a base flare rate for each light curve. We then calculate how often each orbital phase has been observed, also for each light curve separately, to ensure that the different detection thresholds for TESS and Kepler do not bias the base rate. With the phase coverage and base flare rate combined, we can tell how many flares we would expect to see in any given phase bin if flares were randomly distributed. We then aggregate this number of expected flares per time bin into one expected distribution by summing over all available light curves. As a last step, we compare the expected distribution to the observed one using an Anderson-Darling test. We repeat this test with different phase offsets to account for potential biases in sensitivity of the test at different phases.

@@

\subsection{Power of star-planet interaction}
\label{sec:methods:pspi}
We consider two theories for the mechanism behind flaring SPI, the stretch-and-break and the Alfv\'en wing mechanisms. For both of them, we include the case of a magnetized and an unmagnetized planet, for a total of four estimates of the expected power of magnetic SPI.

First, we estimate the power of SPI generated by the stretch-and-break mechanism using the scaling relations in~\cite{lanza2012starplanet} (their Eqn. 45), where $P_{\rm spi,sb}$ is proportional to

\begin{equation}
    P_{\rm spi,sb} \sim B_p^{2/3} R_p^2 B_*^{4/3} R_*^4 v_{rel} a^{-4} .
\end{equation}

@@

\begin{equation}
    P_{\rm spi,sb0} \sim R_p^2 R_*^4 B_*^2 v_{rel}    a^{-4} 
\end{equation}

Third, we estimate the power of SPI generated by the Alfv\'en wing mechanism ~\citep{saur2013magnetic,kavanagh2022radio}, combining Eqns.~8~and~11 in \citet{kavanagh2022radio} to arrive at

\begin{equation}
    P_{\rm spi,aw} \sim B_p^{2/3} R_{p}^2    B_*^{1/3}  v_{rel}^2 a^{-2}  
\end{equation}
where we replace the stellar wind density $\rho_W$ with its proportionality to $a^{-2}$, and stellar wind magnetic field at the planetary radius $B_W$ with the surface field strength $B_* a^{-3}$, which result in the total $a^{-2}$ factor in the above equation.

@@

\begin{equation}
    P_{\rm spi,aw0} \sim R_p^2  B_*  v_{rel}^2  a^{-4}
\end{equation}

Lacking $B_p$, and various additional factors that appear in $P_{\rm spi,sb},P_{\rm spi,sb0},P_{\rm spi,aw}$, and  $P_{\rm spi,aw0}$, such as the angle between the relative velocity and magnetic field of the stellar wind~\citep{kavanagh2022radio}, or the force-free parameter~\citep{lanza2012starplanet}, we restrict ourselves to differential estimates, and normalize all $P_{\rm spi}$ to AU Mic.

@@

Our goal was to measure magnetic SPI as the statistical clustering of flares in phase with the innermost planet's orbit. We searched all star-planet systems that were observed with Kepler and TESS at 1 or 2 min cadence, respectively, for flares. For the star-planet systems in the resulting flare catalog~(Section~\ref{sec:results:catalog}), we confirmed that the orbital phase could be known well enough for the entire observing baseline of the system~(Section~\ref{sec:results:coherence}). We calculated the orbital phases of each flare, and the flare phase distributions for each system~(Section \ref{sec:results:phasedist}). We then applied the custom Anderson-Darling test to assess how much each distribution was different from randomly distributed intrinsic flaring. Comparing the test results with the expected power of magnetic SPI, we found our main result: The clustering of flares in orbital phase tentatively increases with the expected power in all considered scenarios of magnetic SPI~(Section~\ref{sec:results:spi}). However, not all systems with high expected power show signs of flaring SPI, which creates two branches in the distribution of test results. Finally, in Section~\ref{sec:results:individualstars}, we provide additional context for the most interesting systems in our sample, and explain why we had to exclude others from the analysis.

@@

Figure~\ref{fig:adtest_bp} shows that in every scenario we considered (Alfv\'en wing or stretch-and-break mechanism, with magnetized or unmagnetized planets), there are two branches on the high end of expected powers. On one branch, the significance of the measured interaction increases with the expected power $P_{\rm spi}$. On the other branch, no interaction is measured regardless of $P_{\rm spi}$.

Table~\ref{tab:maintable_der} lists the $p$-values for each star-planet system. There is no star-planet system for which we find a $>3\sigma$ ($p<0.0027$) signal of flaring SPI in our sample. However, overall, the significance of the deviation increases with higher expected power $P_{\rm spi}$ of interaction, where additional subscripts denote different scenarios introduced in Section~\ref{sec:methods:pspi}. Table~\ref{tab:maintable_der} also lists the derived parameters required in these scaling laws, i.e., Rossby number $R$o, surface-average magnetic field strength $B$, and relative velocity between the planet and a co-rotating magnetic field $v_{\rm rel}$.

@@

\begin{table*}
\footnotesize
\movetableright=-20mm
\script{paper_main_table.py}
\caption{Flaring star-planet interaction. $Ro$, $B$, and $v_{rel}$, are derived from literature values (Table~\ref{tab:maintable_lit}). $P_{\rm spi,xx}$ stands for the power of stretch-and-break ($sb$) and Alfv\'en wing ($aw$) interaction mechanisms, assuming the planet has a magnetic field strength of $1\,$G. $P_{\rm spi,xx0}$ is the same, but assuming an unmagnetized planet. All powers are normalized to AU Mic. The $p$-value of the Anderson-Darling test is lower when the system shows more flares periodic with the planetary orbit.}
\input{output/table_der_vals.tex}
    \label{tab:maintable_der}

@@

\begin{figure*}[ht!]
    \script{paper_adtest_vs_value_scatterplots.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/PAPER_ADtest_bg_main.png}
        \caption{
            Expected power of magnetic SPI vs. AD test results, assuming the stretch-and-break scenario. While the distribution of $p$-values is consistent with no interaction, all scenarios indicate lower $p$-values only for high expected powers of interaction. See Fig.~\ref{fig:adtest_ro} in the Appendix for the same figure, color-coded by Rossby number.
        }
        \label{fig:adtest_bp}
    \end{centering}
\end{figure*}

\begin{figure}[ht!]
    \script{paper_adtest_vs_value_scatterplots.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/PAPER_ADtest_bg_rest.png}
        \caption{
            Same as Fig.~\ref{fig:adtest_bp}, but for three other scenarios. \textbf{Top panel}: stretch-and-break scenario for an unmagnetized planet. \textbf{Middle panel}: Alfv\'en wing scenario, assuming a magnetized planet. \textbf{Bottom panel}: Alfv\'en wing scenario, assuming an unmagnetized planet. See Fig.~\ref{fig:adtest_ro} in the Appendix for the same figure, color-coded by Rossby number.
        }
        \label{fig:adtest_bp_rest}
    \end{centering}
\end{figure}

\subsection{Individual systems}
\label{sec:results:individualstars}
The properties of star-planet systems in this work are diverse, including both very fast and very slowly rotating host stars; systems with super-Earths, Neptunes, and Hot Jupiters; and spectral types covering the lower main sequence from mid-F to late M.

For our analysis, it is important to know that the observed flares indeed occurred on the planet host star. We therefore excluded some systems due to documented contamination from nearby (bound or co-moving) companions~(Section~\ref{sec:results:individualstars:excluded}). One exception to this rule is Kepler-411, a binary system for which the flare contributions from each component could be separated, but which was excluded for other reasons. We also dropped GJ 1061 from the final sample because its rotation period was unconstrained. Another uncertainty in our analysis is the possibility of additional planets at even shorter orbits than the currently known innermost planet, such as might be the case for Proxima Cen~(see Section~\ref{sec:prominensystems}).

The bulk of the systems in our sample are neither expected to show high power of magnetic SPI, nor do they show any deviation from intrinsic stellar flaring. In Section~\ref{sec:prominensystems}, in the subsample of stars with high expected power, of the order of AU Mic's and above, we first take a closer look at those that seem to follow an increase in measured flaring SPI with increasing expected power $P_{\rm spi}$, that is AU Mic itself, K2-25, TOI-540, and HIP 67522. Then we consider those systems where high power was expected in all scenarios, but not measured, i.e., TAP~26 and KOI-12.
\subsubsection{Excluded targets}
\label{sec:results:individualstars:excluded}

@@

\subsubsection{Prominent systems}
\label{sec:prominensystems}
\paragraph{Proxima Cen}
\label{sec:results:individualstars:proxima}
Although Proxima Cen is part of a triple stellar system, we treat it a single star in this work because of the large angular separation of over 2 deg to its companions, $\alpha$ Cen AB. We do not detect a deviation from intrinsic flaring on Proxima Cen, consistent with numerical models that place Proxima Cen b well outside the sub-Alfv\'enic zone~\citep{alvarado-gomez2020earthlike, kavanagh2021planetinduced, garraffo2022revisiting}. However, this might not entirely rule out flaring SPI in Proxima Cen. The tentatively detected Proxima Cen d, a planet further in, at 0.029 au, or 5 day orbital period~\citep{faria2022candidate, artigau2022linebyline} could still cause flaring SPI. Depending on the phase of the activity cycle, Proxima Cen d may be in the sub-Alfv\'enic regime during Proxima Cen's activity maximum~\citep{alvarado-gomez2020earthlike}. However, the orbital period of the tentative planet is so uncertain at this point that its coherence time~(Section~\ref{sec:results:coherence}) is shorter than the roughly 2yr observing baseline of Proxima Cen, preventing us from accurately measuring the flare phases.

@@

AU Mic is a 16-29 Myr old pre-main sequence M0-M1 dwarf with a strong magnetic field of about $3010\pm220\,$G obtained from Zeeman broadening measurements~\citep{reiners2022magnetism}. The innermost Neptune-sized planet, AU Mic b, could be both inside the sub-Alfv\'enic zone if the star's mass loss rate is relatively low at about 30 times the solar value or lower~\citep{alvarado-gomez2022simulating}. If the mass loss rate is high, exceeding the solar value by a factor of several hundreds, AU Mic b could orbit in the super-Alfv\'enic zone, and become unable to experience any planet-induced flaring. Recently, \citet{klein2022one} found tentative periodicity with the orbital period of AU Mic b in the chromospheric He I D line, which increases if the contribution from flares is included in the calculation. Overall, our marginal signal of flaring SPI in the AU Mic system supports the idea that this young, magnetically active M dwarf system with a close-in Neptune could exhibit planet-induced flaring.
For a detailed discussion of AU Mic and its flaring SPI, we refer to~\cite{ilin2022searching}, where we also estimate that an additional 50–100 days of TESS-like monitoring of AU Mic would yield a $3\sigma$ detection if the marginal signal in our data is real.

K2-25 is a multiplanet system around a fast rotating ($P_{\rm rot}<2\,$d) mid-M dwarf in the Hyades open cluster~\citep[600-800 Myr, ][]{stefansson2020habitable}. Both the expected power of interaction, and the measured deviation, are remarkably similar to the AU Mic system. Interestingly, K2-25 b is in a presumably similar environment to GJ 436 b, but does not show the atmospheric escape in Ly$\alpha$ that the latter famously exhibits, which might be due to high speed stellar winds that suppress atmospheric escape~\citep{rockcliffe2021lya, carolan2020dichotomy, vidotto2020stellar}. Further investigation of magnetic SPI, flaring or otherwise, in this system could add an important constraint on K2-25's wind properties by clarifying whether the planet orbits in- or outside the Alfv\'en radius, especially accounting for its moderately high eccentricity~(see also Section~\ref{sec:discussion:eccentricity}).

@@

\paragraph{TAP 26}
\label{sec:results:individualstars:tap26}
TAP 26 is a weak-line T-Tauri star with a strong magnetic field~\citep{yu2017hot}. \citet{lanza2018closeby} estimate that the system can release more energy in magnetic SPI than other  systems with Hot Jupiters, such as HD 179949, that has previously been detected with chromospheric variability in phase with the planet's orbit~\citep{shkolnik2008nature}. However, for TAP 26 b, the orbital period might not be well-constrained. \citet{yu2017hot} apply several methods to derive $P_{\rm orb}$ from the radial velocity data: While the $10.8\,$d orbit we adopted is the most likely according to \citet{yu2017hot}, a $9.0\,$d orbit is also likely, and some of their applied methods favored a $13.4\,$d period. The absence of interaction could hence be a consequence of uncertain orbital period. However, intermittent interactions and the viewing geometry of TAP 26 could also explain the absence of phase-correlated flaring~(see Sections~\ref{sec:discussion:intermittency} and \ref{sec:discussion:viewing}).

\paragraph{KOI-12}
\label{sec:results:individualstars:koi12}
KOI-12, also known as Kepler-448, is a two-planet system with an eccentric outer planet KOI-12 c, and an inner Warm Jupiter planet KOI-12 b in a $17.8\,$d orbit~\citep{masuda2017eccentric}. With respect to rotation rate, stellar and planetary radius, the KOI-12 system is similar to HIP~67522 and TAP~26 (although for the those young stars the radii are large because the stars are still contracting onto the main sequence). In constrast, KOI-12 is $1.4\pm0.3\,$Gyr old~\citep{bourrier2015sophie}, an F5 subgiant~\citep{frasca2016activity}, which is already evolving off the main sequence. Yet, despite its high expected power of interaction, we detect no excess flaring in phase with KOI-12 b. The reason could be that KOI-12 b is in fact super-Alfv\'enic due to the relatively wide orbit compared to TAP 26 and HIP 67522. However, our estimate of the Alfv\'en radius place KOI-12 b in the sub-Alfv\'enic regime~(see Fig.~\ref{fig:as}), although the scaling law applied there may be inaccurate for post-main sequence stars. Intermittent interaction might play a role here as well, and could be favored if future observations capture episodes during which the interaction is 'on'.

@@

Magnetic SPI can only take place if the planet spends at least a part of its orbit in the sub-Alfv\'enic regime~(see Section~\ref{sec:intro:mspi}). We estimated the average Alfv\'en surface size ($\overline{AS}_{R}$) using the power law established by \citet{Chebly2023}:

@@

The third model uses the gravitational perturbation as a proxy for the interaction~\citep{cuntz2000stellar}, originally derived for SPI. Their Eqn. 1 proposes a gravitational perturbation proportional to the mass ratio between planet and star, and $a^{-3}$.
